ScPy: A SuperCollider Extension for Performing Numerical Computation via Embedded Python

### 2016-08-25 doc/paper.tex
little corrections and fixes

\usepackage{fancyhdr}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{minted}
\usepackage{biblatex}
\usepackage{booktabs}

@@

\providecommand*{\listingautorefname}{Listing}

@@

\section{Introduction}

Although a number of open-source projects which interface other languages with SuperCollider exist
\cite{systemsinterfacingwithsc}~\cite{magnusson2011ixi}~\cite{orlarey2009faust}, they are mostly
SuperCollider clients, which have features to control synths and send other automation messages.
One notable exception is OctaveSC~\cite{octavesc}, which embeds GNU Octave within SuperCollider.
However, it is designed primarily for performing operations on control rate data arrays and its
author notes performance issues which prevent it from being used on audio rate data.

One of the goals of this project is to enable more flexible experimentation with FFT and phase
vocoder based operations. SuperCollider contains a total of 32 built in phase vocoder operations,
which provide the ability to produce many common effects. There are additionally some extra
operations available as user created extensions. However, many conceivable effects are impossible.

@@

\section{Methodology}

During the initial planning stage of the project, we wanted to essentially provide a set of
fundamental operations on spectral data which could be composed together to perform any conceivable
effect. It was also a requirement that these operations would be syntactically concise in order to
enable easy and fast experimentation. Ideally, the system would be able to handle complex effects
in real-time.

Although designing a domain-specific language (DSL)~\cite{van2000domain} for this task was briefly
considered, we eventually decided to use Python for a number of reasons. First, Python includes an
excellent C API which makes embedding it within other languages trivial in comparison to many other
options.  Second, Python's syntax is easily readable and concise --- it clearly adheres to our
goals. Finally, Python has a massive number of community built libraries for performant scientific
and numerical computing. With a few small exceptions, these libraries contained every one of the
fundamental spectral operations we could think of.

@@

\section{Usage}

@@

\section{Implementation}

@@

\begin{table}[H]
    \caption{Type correspondence between SuperCollider and Python}
    \begin{center}
        \begin{tabular}{cc}
            \toprule
            SuperCollider type & Python type \\
            \midrule
            \texttt{Float} & \texttt{float} \\
            (subclass of) \texttt{UGen} & \texttt{ndarray} of \texttt{float} \\
            \texttt{Buffer} & \texttt{ndarray} of \texttt{float} \\
            \texttt{FFT} & \texttt{ndarray} of \texttt{complex} \\
            \texttt{Array} & \texttt{list} \\
            \bottomrule
        \end{tabular}
    \end{center}
\end{table}

>>>>

\usepackage{fancyhdr}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{minted}
\usepackage[style=numeric-comp]{biblatex}
\usepackage{booktabs}

@@

\providecommand*{\listingautorefname}{Listing}
\renewcommand{\sectionautorefname}{Section}

@@

\section{Introduction}\label{sec:introduction}

Although a number of open-source projects which interface other languages with SuperCollider exist
\cite{systemsinterfacingwithsc,magnusson2011ixi,orlarey2009faust}, they are mostly SuperCollider
clients, which have features to control synths and send other automation messages.  One notable
exception is OctaveSC~\cite{octavesc}, which embeds GNU Octave within SuperCollider.  However, it
is designed for performing operations on control rate data arrays and suffers from performance
issues which prevent it from being used on audio rate data.

One of the goals of this project is to enable more flexible experimentation with FFT and phase
vocoder based operations. SuperCollider contains a total of 32 built in phase vocoder operations,
which provide the ability to produce many simple effects. There are additionally some extra
operations available as user created extensions. However, many conceivable effects are impossible.
Some examples of novel effects can be found in~\autoref{sec:usage}.

@@

\section{Methodology}\label{sec:methodology}

During the initial planning stage of the project, we wanted to essentially provide a set of
fundamental operations on spectral data which could be composed together to perform any conceivable
effect. It was also a requirement that these operations would be syntactically concise to enable
easy and fast experimentation. Ideally, the system would be able to handle complex effects in
real-time.

Although designing a domain-specific language (DSL)~\cite{van2000domain} for this task was briefly
considered, we eventually decided to use Python for a number of reasons. First, Python includes an
excellent C API which makes embedding it within other languages trivial in comparison to many other
options.  Second, Python's syntax is easily readable and concise --- it clearly adheres to our
goals. Finally, Python has a massive number of community built libraries for performant scientific
and numerical computing. With a few small exceptions, which are clearly specific to spectral data
such as the phase vocoder, these libraries contained every one of the fundamental operations we
could think of.

@@

\section{Usage}\label{sec:usage}

@@

\section{Implementation}\label{sec:implementation}

@@

\begin{table}[H]
    \caption{Type correspondence between SuperCollider and Python}
    \begin{center}
        \begin{tabular}{ll}
            \toprule
            SuperCollider type & Python type \\
            \midrule
            \texttt{Float} & \texttt{float} \\
            (subclass of) \texttt{UGen} & \texttt{ndarray} of \texttt{float} \\
            \texttt{Buffer} & \texttt{ndarray} of \texttt{float} \\
            \texttt{FFT} & \texttt{ndarray} of \texttt{complex} \\
            \texttt{Array} & \texttt{list} \\
            \bottomrule
        \end{tabular}
    \end{center}
\end{table}
